---
title: "Project 2"
author: "Yusi"
date: "3/8/2020"
output: html_document
---


#IBM HR anaytic employee
```{r}
##install.packages("MASS",dependencies = TRUE, repos = "http://cran.us.r-project.org")
library(tidyverse)
library(MASS)
library(caret)
library(car)
library(randomForest)
library(cowplot)
library(corrplot)
library(leaps)
library(e1071)

```


```{r}
#load the dateset
IBMemployee.df <- read.csv ("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2-data.csv", sep= ",", header= TRUE)
IBMemployee.df

#take a look of the dataset
str(IBMemployee.df)
head(IBMemployee.df)
summary(IBMemployee.df)
#looks like there are three category is not relateble.(EmployeeCount, Over18, Standardhour), we will remove it from dataset.
IBMemployee.df [IBMemployee.df$EmployeeCount!=1,]
IBMemployee.df [IBMemployee.df$Over18 != "Y",]
IBMemployee.df [IBMemployee.df$StandardHours!= 80,]
IBMemp.df<- IBMemployee.df %>% dplyr::select(
  Age,
  Attrition,
  BusinessTravel,
  DailyRate,
  Department,
  DistanceFromHome,
  Education,
  EducationField,
  EnvironmentSatisfaction,
  Gender,
  HourlyRate,
  JobInvolvement,
  JobLevel,
  JobRole,
  JobSatisfaction,
  MaritalStatus,
  MonthlyIncome,
  MonthlyRate,
  NumCompaniesWorked,
  OverTime,
  PercentSalaryHike,
  PerformanceRating,
  RelationshipSatisfaction,
  StockOptionLevel,
  TotalWorkingYears,
  TrainingTimesLastYear,
  WorkLifeBalance,
  YearsAtCompany,
  YearsInCurrentRole,
  YearsSinceLastPromotion,
  YearsWithCurrManager
)
IBMemp.df
summary(IBMemp.df)
#checking missing value
sapply(IBMemp.df, function(x) sum(is.na(x)))
#there is no missing values.

```
```{r}
#changing the class of variable from numeric to factor
IBMemp.df$Education<-factor(IBMemp.df$Education)
IBMemp.df$EnvironmentSatisfaction<- factor(IBMemp.df$EnvironmentSatisfaction)
IBMemp.df$JobInvolvement<-factor(IBMemp.df$JobInvolvement)
IBMemp.df$JobLevel<-factor(IBMemp.df$JobLevel)
IBMemp.df$JobSatisfaction<-factor(IBMemp.df$JobSatisfaction)
IBMemp.df$PerformanceRating<-factor(IBMemp.df$PerformanceRating)
IBMemp.df$RelationshipSatisfaction<- factor(IBMemp.df$RelationshipSatisfaction)
IBMemp.df$StockOptionLevel<-factor(IBMemp.df$StockOptionLevel)
IBMemp.df$WorkLifeBalance<- factor(IBMemp.df$WorkLifeBalance)

```

```{r}
numeric.va<- sapply(IBMemp.df, is.numeric)
corr.matrix <- cor(IBMemp.df[,numeric.va])
corrplot(corr.matrix, main='Correaltion plot for numeric variables', method= "number")
#result of this correlation plot:
#Age variable is correlated with TotalWorkingYears
#TotalWorkingYears correlated with MonthlyIncome
#YearsWithCurrManager also correlated with YearsAtCompany
#YearsWithCurrManger correlated with YearsInCurrentRole
#YearsInCurrentRole correlated with YearsAtCompany

#There is some confusion about monthlyrate,dailyrate, hourlyrateand monthlyincome so I will do a plot and see if they are any important to be discovered.
cor.test(IBMemp.df$MonthlyRate, IBMemp.df$HourlyRate)
#result is r value is -0.01603517, correlation is between -1 to 1. So no evidence to prove there is a correaltion between hourly rate to monthly rate.
cor.test(IBMemp.df$MonthlyRate, IBMemp.df$DailyRate)
#r value is -0.02764077, there is no evidene to prove monthly rate has correlatin to daily rate.
cor.test(IBMemp.df$DailyRate, IBMemp.df$HourlyRate)
#r value is 0.04849597, there is not strong evidence to prove there is correlation between daily rate and hourly rate.
cor.test(IBMemp.df$MonthlyIncome,IBMemp.df$MonthlyRate)
#r value is 0.06459407, there is not strong evidence to prove there is correlation between monthly income to monthly rate
cor.test(IBMemp.df$MonthlyIncome, IBMemp.df$HourlyRate)
#r value is 0.002391151, there is no eivdence show that there is correlation between Monthly income and hourly rate.
cor.test(IBMemp.df$MonthlyIncome, IBMemp.df$DailyRate)
#r value is 0.00008790339, there is no evidence to shwo that there is correlation between monthly income and daily rate.

boxplot(IBMemp.df$MonthlyIncome ~ IBMemp.df$Attrition,
        main = "Monthly Income by Attrition",
        horizontal = T,
        xlab = "Dollars" )
boxplot(IBMemp.df$MonthlyRate ~ IBMemp.df$Attrition,
        main= "Monthly rate by Attrition",
        horizontal = T,
        xlab = "Dollars")
boxplot(IBMemp.df$DailyRate ~ IBMemp.df$Attrition,
        main= "Daily rate by Attrition",
        horizontal = T,
        xlab = "Dollars")
boxplot(IBMemp.df$HourlyRate ~ IBMemp.df$Attrition,
        main= "Hourly rate by Attrition",
        horizontal = T,
        xlab = "Dollars")
pairs(~MonthlyIncome + HourlyRate + DailyRate + MonthlyRate,
            data = IBMemp.df,
            main = "Scatterplot for combination")

str(IBMemp.df)
```
# I do not see a correlation between monthly rate and daily rate and hour rate to monthly income, so these will not be included in the analysis. 

```{r}
IBMemp.df$JobLevel <- as.numeric(IBMemp.df$JobLevel)
cor.test(IBMemp.df$MonthlyIncome, IBMemp.df$JobLevel)
cor.test(IBMemp.df$MonthlyIncome, IBMemp.df$TotalWorkingYears)
```

# correlation between monthly income to job level is 0.9516, between monthly income to total working years = 0.7785.



#Checking the p-value for each variables.
```{r}
IBMemp.lm <-lm(MonthlyIncome ~ ., data = IBMemp.df)
summary(IBMemp.lm)
vif(IBMemp.lm)
```
#run the model, linear regression.
```{r}
trainControl(method = "cv", number = 4) -> train.cv
train(MonthlyIncome ~ .,
  data = IBMemp.df,
  method = "lmStepAIC",
  trControl = train.cv
) -> IBM_salary.step
summary(IBM_salary.step)
IBM_salary.step$results

```

# Monthly income prediction model
```{r}
str(IBMemp.df)
IBMemp.df$Monthlyincomelm <- predict(IBM_salary.step, newdata= IBMemp.df) 
RMSE(IBMemp.df$Monthlyincomelm, IBMemp.df$MonthlyIncome)/mean(IBMemp.df$MonthlyIncome)

```
# RMSE 0.154

# Run the prediction data
```{r}
predictmonthlyincome<- read_excel("/Users/yusi/Downloads/CaseStudy2CompSet No Salary.xlsx")
predictmonthlyincoreame
str(predictmonthlyincome)
head(predictmonthlyincome)
predictmonthlyincome$Education<-factor(predictmonthlyincome$Education)
predictmonthlyincome$EnvironmentSatisfaction<- factor(predictmonthlyincome$EnvironmentSatisfaction)
predictmonthlyincome$JobInvolvement<-factor(predictmonthlyincome$JobInvolvement)
predictmonthlyincome$JobLevel<-factor(predictmonthlyincome$JobLevel)
predictmonthlyincome$JobSatisfaction<-factor(predictmonthlyincome$JobSatisfaction)
predictmonthlyincome$PerformanceRating<-factor(predictmonthlyincome$PerformanceRating)
predictmonthlyincome$RelationshipSatisfaction<- factor(predictmonthlyincome$RelationshipSatisfaction)
predictmonthlyincome$StockOptionLevel<-factor(predictmonthlyincome$StockOptionLevel)
predictmonthlyincome$WorkLifeBalance<- factor(predictmonthlyincome$WorkLifeBalance)
predictmonthlyincome$MonthlyIncomeLM <- predict(IBM_salary.step, newdata = predictmonthlyincome)

predictmonthlyincome
load("workspace.RData")
```

```{r}
boxplot(IBMemp.df$MonthlyIncome ~ IBMemp.df$PerformanceRating)
```
# This is the boxplot for performance rating between excellent as 3 oustanding as 4. It doesnt have much impact on monthly income or attrition. 
#After data cleaning, build the new dataframe. 
```{r}
IBMclean.df <- IBMemp.df %>% dplyr::select(
  MonthlyIncome,
  Attrition,
  Age,
  BusinessTravel,
  Department,
  DistanceFromHome,
  Education,
  EducationField,
  EnvironmentSatisfaction,
  Gender,
  JobInvolvement,
  JobLevel,
  JobRole,
  JobSatisfaction,
  MaritalStatus,
  NumCompaniesWorked,
  OverTime,
  PercentSalaryHike,
  RelationshipSatisfaction,
  StockOptionLevel,
  TotalWorkingYears,
  TrainingTimesLastYear,
  WorkLifeBalance,
  YearsAtCompany,
  YearsInCurrentRole,
  YearsSinceLastPromotion,
  YearsWithCurrManager,
  PerformanceRating
)
```
## Including Plots

#Plot histogram for all numeric variables

```{r}
IBMclean.df %>% keep(is.numeric) %>% gather %>% ggplot(aes(x= value))+ facet_wrap(~key, scales = "free")+
  geom_histogram()
                     
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Age is more or less normally distributed.
# Distance from Home, Num of Companies worked and Total working Years is right skewed and should be transformed to curb the skewness.
# MonthlyIncome, PercentSalatyHike, Yearatcompany, Yearsincurrentrole, years since last promotion, years with current managers all are right skewed and need transformed to meet the assumption.


#Employee Personal Demographics 
```{r}
p1<- IBMclean.df %>%
  group_by(Gender) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(Gender), y = counts)) + geom_bar(stat = 'identity', fill = "red") + ggtitle("Gender") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 900))

p2<- IBMclean.df %>%
  group_by(Education) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(Education), y = counts)) + geom_bar(stat = 'identity', fill = "blue") + ggtitle("Education") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 650))

p3 <- IBMclean.df %>%
  group_by(EducationField) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(EducationField), y = counts)) + geom_bar(stat = 'identity', fill = "yellow") + ggtitle("Education Field") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 650))

p4 <- IBMclean.df %>%
  group_by(MaritalStatus) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(MaritalStatus), y = counts)) + geom_bar(stat = 'identity', fill = "green")+ ggtitle("Marital Status") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 750))

p5 <- IBMclean.df %>%
  group_by(RelationshipSatisfaction) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(RelationshipSatisfaction), y = counts)) + geom_bar(stat = 'identity', fill = "coral1") + ggtitle("Relationship Satisfaction") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())+ scale_y_continuous(limits = c(0, 500))

p6 <- IBMclean.df %>%
  group_by(WorkLifeBalance) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(WorkLifeBalance), y = counts)) + geom_bar(stat = 'identity', fill = "purple")+ ggtitle("Work Life Balance") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 950))

grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 2, ncol = 3)

```
# More than 50% of Male employees
#Education: 1 'Below College 2 'College' 3 'Bachelor' 4 'Master' 5 'Doctor'; majority employees has bachelor degree, second one is master degree.
# More than 75% of employees are from life science or medical education backgroud. 
# 45% of employees are married. 
# Realtionship Satisfaction: 1 'Low' 2 'Medium' 3 'High' 4 'Very High'. 
# 1 'Bad' 2 'Good' 3 'Better' 4 'Best'; looks like majority of employees think they have a better work like balance in IBM.

# Work Demographics 
```{r}
p1 <- IBMclean.df %>%
  group_by(BusinessTravel) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(BusinessTravel), y = counts)) + geom_bar(stat = 'identity', fill = "red") + ggtitle("Business Travel") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10,angle = 45, hjust = 1),axis.title.x=element_blank())+ scale_y_continuous(limits = c(0, 1100))



p2 <- IBMclean.df %>%
  group_by(EnvironmentSatisfaction) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(EnvironmentSatisfaction), y = counts)) + geom_bar(stat = 'identity', fill = "green") + ggtitle("Environment Satisfaction") + geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 500))

p3 <- IBMclean.df %>%
  group_by(JobInvolvement) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(JobInvolvement), y = counts)) + geom_bar(stat = 'identity', fill = "blue") + ggtitle("Job Involvement") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 900))


p4 <- IBMclean.df %>%
  group_by(JobSatisfaction) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(JobSatisfaction), y = counts)) + geom_bar(stat = 'identity', fill = "purple") + ggtitle("Job Satisfaction") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 500))

p5 <- IBMclean.df %>%
  group_by(OverTime) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(OverTime), y = counts)) + geom_bar(stat = 'identity', fill = "coral1") + ggtitle("Over Time") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) + scale_y_continuous(limits = c(0, 1100))


p6 <- IBMclean.df %>%
  group_by(Department) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(Department), y = counts)) + geom_bar(stat = 'identity', fill = "grey") + ggtitle("Department") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ theme(plot.title = element_text(size =10),axis.text.x = element_text(size = 7, angle = 45, hjust = 1),axis.title.x=element_blank())

p7 <- IBMclean.df %>%
  group_by(JobRole) %>%
  summarise(counts = n()) %>%
  ggplot(aes(x = as.factor(JobRole), y = counts)) + geom_bar(stat = 'identity', fill = "red") + ggtitle("Job Role") +geom_text(aes(label=counts), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25) + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())
grid.arrange(p1,p2,p3,p4,p5,p6,p7,nrow = 2)
```
#Now let's take a look of the realtionship to attrtion.
```{r}
p1 <- IBMclean.df %>%
  ggplot(aes(x = Age, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Age") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- IBMclean.df %>%
  ggplot(aes(x = DistanceFromHome, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Distance From Home")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- IBMclean.df %>%
  ggplot(aes(x = NumCompaniesWorked, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Number of Companies")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p4 <- IBMclean.df %>%
  ggplot(aes(x = TotalWorkingYears, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Total Working Years")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)

```
# Younger employees between 25- 35 years old have a hight attrition rate.
# Attrtion rate is lower when distance from home is within 10 miles.
# Attrtion rate is higher when employees have total working years between 0.7 years. 

```{r}

p1 <- IBMclean.df %>%
  ggplot(aes(x = MonthlyIncome, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Monthly Income") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p2 <- IBMclean.df %>%
  ggplot(aes(x = PercentSalaryHike, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Percentage Salary Hike") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p3 <- IBMclean.df %>%
  ggplot(aes(x = YearsAtCompany, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years At Company") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p4 <- IBMclean.df %>%
  ggplot(aes(x = YearsInCurrentRole, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years in Current Role") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p5 <- IBMclean.df %>%
  ggplot(aes(x = YearsSinceLastPromotion, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years Since Last Promotion") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p6 <- IBMclean.df %>%
  ggplot(aes(x = YearsWithCurrManager, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years With Current Manager") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


grid.arrange(p1, p2, p3, p4, p5, p6 , nrow = 3, ncol = 2)
```
# feature selection
```{r}
p1 <- IBMclean.df %>%
  group_by(BusinessTravel) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = BusinessTravel, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Business Travel") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 30))
p2 <- IBMclean.df %>%
  group_by(EnvironmentSatisfaction) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = EnvironmentSatisfaction, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Environment Satisfaction") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 30))

p3 <- IBMclean.df %>%
  group_by(JobInvolvement) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobInvolvement, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Job Involvement") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 40))

p4 <- IBMclean.df %>%
  group_by(JobSatisfaction) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobSatisfaction, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Job Satisfaction") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 30))

p5 <- IBMclean.df %>%
  group_by(OverTime) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = OverTime, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Over Time") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 35))

p6 <- IBMclean.df %>%
  group_by(Department) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = Department, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Department") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)

p7 <- IBMclean.df %>%
  group_by(JobRole) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobRole, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "Blue") + ggtitle("Attrition Rate VS Job Role") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank()) +geom_text(aes(label=attrition_rate), size = 2.5, position=position_dodge(width=0.2), vjust=-0.25)

grid.arrange(p1, p2, p3, p4, p5, p6, p7, nrow = 2, ncol = 4)
```
# Attrtion rate is higher with employees who travel more frequently; also with low envirnment satisfaction, low job satisfaction and low job involvement; higher with over time employees, and sales department. 

#According to the histogram, we need to transform the data to be normaly distributed. 
```{r}
IBMfinal<- IBMclean.df %>% dplyr::select(
  MonthlyIncome,
  Attrition,
  Age,
  BusinessTravel,
  Department,
  DistanceFromHome,
  Education,
  EducationField,
  EnvironmentSatisfaction,
  JobInvolvement,
  JobLevel,
  JobRole,
  JobSatisfaction,
  MaritalStatus,
  NumCompaniesWorked,
  OverTime,
  PercentSalaryHike,
  RelationshipSatisfaction,
  StockOptionLevel,
  TotalWorkingYears,
  TrainingTimesLastYear,
  WorkLifeBalance,
  YearsAtCompany,
  YearsInCurrentRole,
  YearsSinceLastPromotion,
  YearsWithCurrManager,
)

```

```{r}
Train <- createDataPartition(IBMfinal$Attrition, p=0.8, list=FALSE)
training <- IBMfinal[ Train, ]
testing <- IBMfinal[ -Train, ]
prop.table(table(IBMfinal$Attrition))
```


```{r}
train.knn <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 30,
  summaryFunction = twoClassSummary,
  classProbs = TRUE
)
# KNN on Training Set
IBMattrition.knn <- train(
  Attrition ~ .,
  data = training,
  method = "knn",
  metric = "Spec",
  trControl = train.knn,
  preProcess = c("center","scale"),
  tuneLength = 8
)
# Adding predictions to Test Data
predict(IBMattrition.knn, newdata = testing ) -> testing$Attrition_KNN
# creating confusion matrix
confusionMatrix(
  table(testing$Attrition, testing$Attrition_KNN )
)
```


```{r}
IBMpred.df<- read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2CompSet%20No%20Attrition.csv", sep = ",", header = TRUE)
head(IBMpred.df)
str(IBMpred.df)
IBMpred.df$Education<-factor(IBMpred.df$Education)
IBMpred.df$EnvironmentSatisfaction<- factor(IBMpred.df$EnvironmentSatisfaction)
IBMpred.df$JobInvolvement<-factor(IBMpred.df$JobInvolvement)
IBMpred.df$JobLevel<-factor(IBMpred.df$JobLevel)
IBMpred.df$JobSatisfaction<-factor(IBMpred.df$JobSatisfaction)
IBMpred.df$PerformanceRating<-factor(IBMpred.df$PerformanceRating)
IBMpred.df$RelationshipSatisfaction<- factor(IBMpred.df$RelationshipSatisfaction)
IBMpred.df$StockOptionLevel<-factor(IBMpred.df$StockOptionLevel)
IBMpred.df$WorkLifeBalance<- factor(IBMpred.df$WorkLifeBalance)

IBMpred.df$Attrition_KNN <- predict(attrition.knn, IBMpred.df)

IBMpred.df
```



